---
title: "P8105 Data Science I Homework 4"
author: Olivia Wang (hw2852)
output: github_document
date: "2022-11-16"
---

In preparation for the problems below, we will load the following libraries: 

```{r load_libraries}
library(tidyverse)
library(readxl)
library(dplyr)
```

# Problem 1



# Problem 2

Let us begin by importing the CSV file containing _Washington Post's_ homicide data downloaded from GitHub, and applying the `clean_names` function. We can then apply the `skim` function to generate a brief data summary. 

```{r}
homicide_data = 
  read_csv("./homicide-data.csv") %>% 
  janitor::clean_names()

skimr::skim(homicide_data)
```

The _Washington Post's_ raw homicide data contains __`r nrow(homicide_data)` rows__ and __`r ncol(homicide_data)` columns__. Each row corresponds to one homicide case, and the following information regarding the homicide are found in each column:

* Case ID code
* Reported date
* Victim identifiers (name, age, race and sex)
* Location details (state, city, latitude and longitude)
* Disposition

Based on the output generated above, we can see that there are no missing values for all variables except for `latitude` and `longitude`. The `latitude` variable has __`r sum(is.na(homicide_data$lat))` missing values__ and the `longitude` variable has __`r sum(is.na(homicide_data$lon))` missing values__. 


# Problem 3
